Self-test prime dimensional EPR pairs with constant alphabet

### 2019-02-06 Draft/draft.tex
Honghao: start rewriting section 5.1

We use some basic number theory in our work. A primitive root of a prime number
$d$ is an integer $r$ such that $r \pmod{d}$ has multiplicative order $d-1$. Equivalently,
$r$ is the generator of the multiplicative group of integers modulo $d$, $\Z_d^\times$ where
$\Z_d^\times = \{1,2\dots d-1\}$. When we say $d$ has primitive root $r$, we always mean
that $r$ is the smallest primitive root.  \carl{I'm not clear on what that last sentence means.  You just defined what a primitive root
is --- are you changing the definition now?}

@@

%-----------------------------------------------------------------
\subsection{The correlation $C(\dr{r})$}
%-----------------------------------------------------------------
Recall that the linear system game $\LS_r$ has $n_r$ variables and $m_r$ equations.
In game $\G(\dr{r})$, 
Alice receives $x \in \{1,\dots m_r+3 \}$ and Bob receives
$y \in \{1,\dots,n_r+1\}$. We follow the previous structure by first give intuition about how they should
behave in this game. The correlation can be easily extracted from the behaviour list below.
\begin{itemize}
	\item When $x \in \{1,\dots m_r\}$ and $y \in \{1, \dots n_r\}$, they should win the 
	linear system game $\LS_r$ perfectly;
	\item when $x \in \{m_r+1, m_r+2, m_r+3\}$ and $y \in \{1, 2, n_r+1\}$, they should follow the
	optimal correlation of the test $\CHSH_X$, where 
	\begin{align}
		&\ast_A = m_r+1, \quad 0_A = m_r+2,\quad 1_A = m_r+3,\\
		&\ast_B = n_r+1,\quad 0_B = 1, \quad 1_A = 2,
	\end{align}
	are the inputs for the game $\CHSH_X$
	(The intuition behind is that $B_1B_2 = X$.);
	\item otherwise, their behaviour is irrelevant.
\end{itemize}
Note that the dimension $\dr{r}$ is defined in the rules of $\CHSH_X$.

\begin{proposition}
	\label{prop:realize}
	The correlation $C(\dr{r})$ can be realized by a quantum strategy.
\end{proposition}
We prove this proposition by giving the ideal strategy.
\begin{proof}
The construction start with the ideal strategy for the weighted $\CHSH_X$ test.
We choose $\ket{\psi} = \frac{1}{\sqrt{d-1}}\sum_{i=1}^{d-1} \ket{ii}$ and $V = \spn(\ket{1} ,\ket{d-1})$.
Alice's projectors and observables for the weighted $\CHSH_X$ test are
	&A_\ast^\diamond = \ketbra{1}{1} + \ketbra{d-1}{d-1}, &&A_\ast^\perp = \1 - A_\ast^\diamond, \\
	&A_1 = A_1^0-A_1^1 = \ketbra{1}{1} - \ketbra{d-1}{d-1}, &&A_1^\perp = A_\ast^\perp,\\
	&A_2 = A_2^0-A_2^1 = \ketbra{1}{d-1} + \ketbra{d-1}{1} ,&&A_2^\perp = A_\ast^\perp.
Bob's observables restricted to the subspace $V$ is
	B_1|_V = \cos(\frac{\pi}{d})(\ketbra{1}{1} - \ketbra{d-1}{d-1}) - \sin(\frac{\pi}{d})(\ketbra{1}{d-1} + \ketbra{d-1}{1}),\\
	B_2|_V = \cos(\frac{\pi}{d})(\ketbra{1}{1} - \ketbra{d-1}{d-1}) +\sin(\frac{\pi}{d})(\ketbra{1}{d-1} + \ketbra{d-1}{1}).
To make sure operator $B_2$ commutes with the operator $U$ that we will construct later, we pick a 
special basis of $X =B_1B_2$, which is
	\ket{x_1} = \frac{-1}{\sqrt{2}}(\ket{1} + i\ket{d-1}), &&
	\ket{x_{d-1}} = \frac{-e^{i\pi/d}}{\sqrt{2}}(\ket{1} - i\ket{d-1}).
It can be checked that $X\ket{x_1} = \omega_d \ket{x_1}$ and $X\ket{x_{d-1}} = \omega_d^{d-1} \ket{x_{d-1}}$.
In this basis, $B_1$ and $B_2$ are expressed as 
	&B_1|_V = \omega_d\ketbra{x_1}{x_{d-1}} + \omega_d^{-1} \ketbra{x_{d-1}}{x_1},\\
	&B_2|_V =\ketbra{x_1}{x_{d-1}}  + \ketbra{x_{d-1}}{x_1}.

@@

the observables $B_1$ and $B_2$ are extended from $B_1|_V$ and $B_2|_V$ as follows

@@

	B_1 &= \sum_{k=1}^{(d-1)/2}\left( \omega_d^k\ketbra{x_k}{x_{d-k}} + \omega_d^{-k}\ketbra{x_{d-k}}{x_k}\right)\\
	B_2 &= \sum_{k=1}^{(d-1)/2}\left(\ketbra{x_{d-k}}{x_k} + \ketbra{x_k}{x_{d-k}}\right).

@@

It can be checked that $B_1B_2 = \sum_{i=1}^{d-1} \omega_d^i \ketbra{x_i}{x_i}$ and the operator $U$ is

@@

	U = \sum_{i=1}^{d-1} \ketbra{x_{i/r \pmod{d}}}{x_i}.

@@

the commutativity between $U$ and $B_2$, which can be verified as follows,

@@

The decomposition of $U$ into $B_3$ and $B_4$ is similar to the decomposition of $X$.
Suppose the eigen-decomposition of $U$ is

@@

	U = \sum_{i=0}^{d-1} \omega_{d-1}^i \ketbra{u_i}{u_i},

@@

then the decomposition is the following 

@@

	B_3 =& \ketbra{u_0}{u_0} +\omega_{d-1}^{(d-1)/2}\ketbra{u_{(d-1)/2}}{u_{(d-1)/2}} + \sum_{k=1}^{(d-3)/2}\left( \omega_d^k\ketbra{u_k}{u_{d-1-k}} + \omega_d^{-k}\ketbra{x_{d-1-k}}{x_k}\right)\\
	B_4 &= \ketbra{u_0}{u_0} +\ketbra{u_{(d-1)/2}}{u_{(d-1)/2}} + \sum_{k=1}^{(d-3)/2}\left(\ketbra{u_{d-1-k}}{u_k} + \ketbra{u_k}{u_{d-1-k}}\right).

@@

In the next step of embedding, for each $x_i$, we find $y_i$ ,$z_i$, $w_i$ and $f$ such that
$x_i = y_iz_i = fw_i$. We demonstrate how this splitting is done by pick $x_1$ as an example, 
which is mapped to $B_1$ in the previous step. Now we map $x_1$ to $B_1'$ which is
\begin{align}
	B_1' = \begin{pmatrix}
	B_1 & 0 \\
	0 & B_1
	\end{pmatrix},
\end{align}
then $y_1$, $z_1$, $w_1$ and $f$ are mapped to
\begin{align}
y_1 \to 
\begin{pmatrix}
B_1 & 0\\
0 & \1
\end{pmatrix},
&&
z_1 \to
\begin{pmatrix}
\1 & 0\\
0 & B_1
\end{pmatrix},
\\
w_1 \to 
\begin{pmatrix}
0 & B_1\\
B_1 & 0
\end{pmatrix},
&&
f \to
\begin{pmatrix}
0 & \1\\
\1 & 0
\end{pmatrix}.
\end{align}

>>>>

We use some basic number theory in our work. A primitive root of a prime number
$d$ is an integer $r$ such that $r \pmod{d}$ has multiplicative order $d-1$. Equivalently,
$r$ is the generator of the multiplicative group of integers modulo $d$, $\Z_d^\times$ where
$\Z_d^\times = \{1,2\dots d-1\}$. \textcolor{red}{When the prime number $d$ has multiple primitive roots,
we always work with the smallest one.}
\carl{I'm not clear on what that last sentence means.  You just defined what a primitive root
is --- are you changing the definition now?}

@@

%-----------------------------------------------------------------
\subsection{The correlation $C(\dr{r})$}
%-----------------------------------------------------------------
Recall that the linear system game $\LS_r$ has $n_r$ variables and $m_r$ equations.
In game $\G(\dr{r})$, 
Alice receives $x \in \{1,\dots m_r+3 \}$ and Bob receives
$y \in \{1,\dots,n_r+1\}$. We follow the previous structure by first give intuition about how they should
behave in this game. The correlation can be easily extracted from the behaviour list below.
\begin{itemize}
	\item When $x \in \{1,\dots m_r\}$ and $y \in \{1, \dots n_r\}$, they should win the 
	linear system game $\LS_r$ perfectly;
	\item when $x \in \{m_r+1, m_r+2, m_r+3\}$ and $y \in \{1, 2, n_r+1\}$, they should follow the
	optimal correlation of the test $\CHSH_X$, where 
	\begin{align}
		&\ast_A = m_r+1, \quad 1_A = m_r+2,\quad 2_A = m_r+3,\\
		&\ast_B = n_r+1,\quad 1_B = 1, \quad 2_A = 2,
	\end{align}
	are the inputs for the game $\CHSH_X$
	(The intuition behind is that $B_1B_2 = X$.);
	\item otherwise, their behaviour is irrelevant.
\end{itemize}
Note that the dimension $\dr{r}$ is defined in the rules of $\CHSH_X$.

\begin{proposition}
	\label{prop:realize}
	The correlation $C(\dr{r})$ can be realized by a quantum strategy.
\end{proposition}
We prove this proposition by giving the ideal strategy.
\begin{proof}
We give the ideal strategy for the extended weighted $\CHSH_X$ test first and then based on it we give a representation of 
group $\Gamma(\LS)$, $\Psi$, which can be converted to an ideal strategy for $\LS$. 
The ideal strategy for $C(\dr{r})$ is the combination of the two ideal strategies for the two tests.

\textbf{The ideal strategy for the weighted $\CHSH_X$ test}.
Alice and Bob share state $\ket{\psi} = \frac{1}{\sqrt{d-1}}\sum_{i=1}^{d-1} \ket{ii}$ and choose the special subspace to be 
$V = \spn(\ket{1} ,\ket{d-1})$.
Alice's projectors and observables for the weighted $\CHSH_X$ test are
\begin{align*}
	&A_0^\diamond = \ketbra{1}{1} + \ketbra{d-1}{d-1}, &&A_0^\perp = \1 - A_\ast^\diamond, \\
	&A_{m_r+2} = A_{m_r+2}^0-A_{m_r+2}^1 = \ketbra{1}{1} - \ketbra{d-1}{d-1}, &&A_{m_r+2}^\perp = A_\ast^\perp,\\
	&A_{m_r+3} = A_{m_r+3}^0-A_{m_r+3}^1 = \ketbra{1}{d-1} + \ketbra{d-1}{1} ,&&A_{m_r+3}^\perp = A_\ast^\perp.
\end{align*}
Recall that in the extended weighted $\CHSH_X$ test, what matters is the 
Bob's observables restricted to the subspace $V$, which are given by
\begin{align*}
	B_1|_V = \cos(\frac{\pi}{d})(\ketbra{1}{1} - \ketbra{d-1}{d-1}) - \sin(\frac{\pi}{d})(\ketbra{1}{d-1} + \ketbra{d-1}{1}),\\
	B_2|_V = \cos(\frac{\pi}{d})(\ketbra{1}{1} - \ketbra{d-1}{d-1}) +\sin(\frac{\pi}{d})(\ketbra{1}{d-1} + \ketbra{d-1}{1}).
\end{align*}
To make sure operator $B_2$ commutes with the operator $U = \Psi(x_3x_4) = \Psi(u)$ that we will construct later, we pick a 
special basis of $X|_V =B_1|_VB_2|_V$, which is
\begin{align}
	\ket{x_1} = \frac{-1}{\sqrt{2}}(\ket{1} + i\ket{d-1}), &&
	\ket{x_{d-1}} = \frac{-e^{i\pi/d}}{\sqrt{2}}(\ket{1} - i\ket{d-1}).
\end{align}
It can be checked that $X\ket{x_1} = \omega_d \ket{x_1}$ and $X\ket{x_{d-1}} = \omega_d^{d-1} \ket{x_{d-1}}$.
In this basis, $B_1|_V$ and $B_2|_V$ are expressed as 
\begin{align}
	&B_1|_V = \omega_d\ketbra{x_1}{x_{d-1}} + \omega_d^{-1} \ketbra{x_{d-1}}{x_1},\\
	&B_2|_V =\ketbra{x_1}{x_{d-1}}  + \ketbra{x_{d-1}}{x_1}.
\end{align}

\textbf{The ideal strategy for $\LS_r$}.
We denote the number of generators labelled by $x_i$ in $\LS_r$ by $n_r(x)$.
In order to construct a representation of $\Gamma(\LS_r)$, we first construct a representation, $\Psi_{G(x)}$ for the group 
\begin{align}
	G(x) = \ip{ \{x_i\}_{i=1}^{n_r(x)}: x_i^2 = e, x_i x_j x_i = x_k \text{ for all } (i,j,k) \in C_r},
\end{align}
where details of $C_r$ can be found in the appendices. 
Note that the relation $uxu^{-1} = x^2$ is embedded in the conjugation relations in $C_r$.
Building on $\Psi_{G(x)}$, we can construct
the representation $\Psi_{\Gamma(\LS_r)}$ for the solution group $\Gamma(\LS_r)$.

@@

the observables $A_1 = B_1 = \Psi_{G(x)}(x_1)$ and $A_2 = B_2 = \Psi_{G(x)}(x_2)$ are extended from $B_1|_V$ and $B_2|_V$ as follows

@@

	A_1 = B_1 = \Psi_{G(x)}(x_1) &= \sum_{k=1}^{(d-1)/2}\left( \omega_d^k\ketbra{x_k}{x_{d-k}} + \omega_d^{-k}\ketbra{x_{d-k}}{x_k}\right)\\
	A_2 = B_2 = \Psi_{G(x)}(x_2) &= \sum_{k=1}^{(d-1)/2}\left(\ketbra{x_{d-k}}{x_k} + \ketbra{x_k}{x_{d-k}}\right).

@@

It can be checked that $X = \Psi_{G(x)}(x_1x_2) = \sum_{i=1}^{d-1} \omega_d^i \ketbra{x_i}{x_i}$ 
The relation $uxu^{-1} = x^r$ and $X$ determine that 

@@

	U =\Psi_{G(x)}(u) = \Psi_{G(x)}(x_3x_4) =  \sum_{i=1}^{d-1} \ketbra{x_{i/r \pmod{d}}}{x_i}.

@@

the commutativity between $U$ and $B_2$, so we prove the commutation relation first and
then give $\Psi_{G(x)}(x_3)$ and $\Psi_{G(x)}(x_4)$.
The commutation relation can be checked as follows,

@@

The decomposition of $U$ into $\Psi_{G(x)}(x_3)$ and $\Psi_{G(x)}(x_4)$ is similar to the decomposition of $X$.
Since $u$ has order $d-1$, we can find an eigenbasis $\{ \ket{u_i} \}_{i=0}^{d-2}$ such that the eigen-decomposition of $U$ is

@@

	U = \sum_{i=0}^{d-2} \omega_{d-1}^i \ketbra{u_i}{u_i},

@@

then $\Psi_{G(x)}(x_3)$ and $\Psi_{G(x)}(x_4)$ are  
	A_3 = B_3 =\Psi_{G(x)}(x_3) = &\ketbra{u_0}{u_0} +\omega_{d-1}^{(d-1)/2}\ketbra{u_{(d-1)/2}}{u_{(d-1)/2}}\\ + 
	&\sum_{k=1}^{(d-3)/2}\left( \omega_{d-1}^k\ketbra{u_k}{u_{d-1-k}} + \omega_{d-1}^{-k}\ketbra{x_{d-1-k}}{x_k}\right)\\ 
	A_4 = B_4  =\Psi_{G(x)}(x_4) = &\ketbra{u_0}{u_0} +\ketbra{u_{(d-1)/2}}{u_{(d-1)/2}} \\+
	 &\sum_{k=1}^{(d-3)/2}\left(\ketbra{u_{d-1-k}}{u_k} + \ketbra{u_k}{u_{d-1-k}}\right).
The conjugation relations in $C_r$ determines the representations of 
the rest of the $x_i$'s from $x_1,x_2,x_3$ and $x_4$. 
For example, if $x_5 = x_1x_2x_1 \in C_r$, then

@@

	\Psi_{G(x)}(x_5) = \Psi_{G(x)}(x_1)\Psi_{G(x)}(x_2) \Psi_{G(x)}(x_1).

@@

Next, we consider the representation $\Psi_{G(x,y,z,w)}$.
In the next step of embedding, for each $x_i$, we find $y_i$ ,$z_i$, $w_i$ and $f$ such that
$x_i = y_iz_i = fw_i$. We demonstrate how this splitting is done by pick $x_1$ as an example, 
which is mapped to $B_1$ in the previous step. Now we map $x_1$ to $B_1'$ which is
\begin{align}
	B_1' = \begin{pmatrix}
	B_1 & 0 \\
	0 & B_1
	\end{pmatrix},
\end{align}
then $y_1$, $z_1$, $w_1$ and $f$ are mapped to
\begin{align}
y_1 \to 
\begin{pmatrix}
B_1 & 0\\
0 & \1
\end{pmatrix},
&&
z_1 \to
\begin{pmatrix}
\1 & 0\\
0 & B_1
\end{pmatrix},
\\
w_1 \to 
\begin{pmatrix}
0 & B_1\\
B_1 & 0
\end{pmatrix},
&&
f \to
\begin{pmatrix}
0 & \1\\
\1 & 0
\end{pmatrix}.
\end{align}
